---
title: "City Tree Analyses"
author: "Joey Hulbert"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: true
    toc_depth: 3
---

|            |            |            |            |            |
|:----------:|:----------:|:----------:|:----------:|:----------:|
|[Welcome](https://jmhulbert.github.io/redhot)|[Data](https://jmhulbert.github.io/redhot/data)|[Analyses](https://jmhulbert.github.io/redhot/analyses)|[Discussion](https://jmhulbert.github.io/redhot/discussion)|[Maps](https://jmhulbert.github.io/redhot/maps)
|             |           |            |            |            |

|            |            |            | Analyses   |            |            |            |
|:----------:|:----------:|:----------:|:----------:|:----------:|:----------:|:----------:|
|[Urban Heat](https://jmhulbert.github.io/redhot/analyses/heat/)|[All Areas](https://jmhulbert.github.io/redhot/analyses/heat/all-areas)|[Portland](https://jmhulbert.github.io/redhot/analyses/heat/portland)|[Tacoma](https://jmhulbert.github.io/redhot/analyses/heat/tacoma)|[King County](https://jmhulbert.github.io/redhot/analyses/heat/king-county)|[iNaturalist](https://jmhulbert.github.io/redhot/analyses/inaturalist/)|[City Inventories]((https://jmhulbert.github.io/redhot/analyses/city-trees/))|
|             |           |            |            |            |            |            |

# CITY INVENTORIES

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(gghalves)
library(patchwork)
library(scales)
library(ggmap)
library(osmdata)
```


# Purpose

# Data Wrangling

## Import Data

```{r}
portland.parks <- read.csv('https://raw.githubusercontent.com/jmhulbert/redhot/main/data/Portland.Parks.Trees.AF.csv')
portland.streets <- read.csv('https://raw.githubusercontent.com/jmhulbert/redhot/main/data/Portland.Street.Trees.AF.csv')
seattle.combined <- read.csv('https://raw.githubusercontent.com/jmhulbert/redhot/main/data/Seattle.Trees.AF.csv')
```

## Tidy Data

```{r}
portland.parks <- portland.parks[c(3,4,6,7,8,14,35,36,40,42:45)]
portland.parks <- portland.parks %>% mutate(city.space="Park")

portland.streets <- portland.streets[c(3,4,5,9,10:19)]
portland.streets <- portland.streets %>% mutate(city.space="Street") %>% rename(DBH=DIAMETER) %>% rename(Functional=FUNCTIONAL)

portland.combined <- full_join(portland.streets,portland.parks,by = join_by(OBJECTID, Condition, DN_AF1, lat, lon, DBH,city.space,Functional)) %>% mutate(city="Portland") %>% mutate(DN_AFc=DN_AF1)

portland.combined$DN_AF1 <- ((portland.combined$DN_AF1 * 1.8) +32)

```

```{r}
seattle.combined <- seattle.combined[c(1:8,11:14,17:19)] %>% mutate(city="Seattle") %>% rename(GlobalID=GLOBALID)

seattle.streets <- seattle.combined %>% filter(Source.Department=="Seattle Department of Transportation") %>% mutate(city.space="Street")
seattle.parks <- seattle.combined %>% filter(Source.Department=="Seattle Parks and Recreation") %>% mutate(city.space="Park")

seattle.combined.back <- rbind(seattle.streets,seattle.parks)
```


# Categorize Trees

Portland Categories - Dead Tops
Protect from development: <85 
Extra care: 85-88.5
Most concern: >88.5

King County Categories - Dead Tops
Protect from development: <85 
Extra care: 85-87.8
Most concern: >87.8

```{r}
portland.protect <- portland.combined %>% filter(DN_AF1<85) %>% mutate(Category="Protect") %>% droplevels()
portland.care <- portland.combined %>% filter(DN_AF1>85 & DN_AF1<88.5) %>% mutate(Category="Extra Care") %>% droplevels()
portland.concern <- portland.combined %>% filter(DN_AF1>88.5)%>% mutate(Category="Most Concern") %>% droplevels()
portland.combined.cat <- bind_rows(list(portland.protect,portland.care,portland.concern))
```

```{r}
seattle.protect <- seattle.combined.back %>% filter(DN_AF1<85) %>% mutate(Category="Protect") %>% droplevels()
seattle.care <- seattle.combined.back %>% filter(DN_AF1>85 & DN_AF1<88.5) %>% mutate(Category="Extra Care") %>% droplevels()
seattle.concern <- seattle.combined.back %>% filter(DN_AF1>88.5)%>% mutate(Category="Most Concern") %>% droplevels()
seattle.combined.cat <- bind_rows(list(seattle.protect,seattle.care,seattle.concern))
```

87 trees in Seattle did not have UHI data. 1 tree in Portland did not have UHI data. 


## Join City Data

```{r}
cities <- full_join(seattle.combined.cat,portland.combined.cat,by = join_by(OBJECTID, DBH, DN_AF1, lat, lon, city, city.space,GlobalID,Category))
```

# Summarize Data

```{r}
summary <- cities %>% group_by(city,city.space,Category) %>% summarize(n=n(),meanAF=mean(DN_AF1,na.rm=TRUE),n_AF_na=sum(is.na(DN_AF1)))
summary
```

```{r}
ggplot(summary,aes(n,city.space,fill=Category))+geom_col(position=position_dodge(),alpha=0.7) +facet_wrap(~city) +theme_bw() + scale_fill_manual(name="Category",values=c("#7fcdbb","#fe9929","#DDA0DD")) +labs(x="Number of Trees",y="Tree Inventory")
```



```{r}
summary <- cities %>% group_by(city,Category) %>% summarize(n=n(),meanAF=mean(DN_AF1,na.rm=TRUE),n_AF_na=sum(is.na(DN_AF1)))
summary
```


# Portland

```{r}
summary <- portland.combined.cat %>% group_by(city.space,Category) %>% summarize(n=n(),meanAF=mean(DN_AF1,na.rm=TRUE),n_AF_na=sum(is.na(DN_AF1)))
summary
```

```{r}
ggplot(summary,aes(n,city.space,fill=Category))+geom_col(position=position_dodge(),alpha=0.7) +theme_bw() + scale_fill_manual(name="Category",values=c("#7fcdbb","#fe9929","#DDA0DD")) +labs(x="Number of Trees",y="Tree Inventory")
```


## Portland Map

```{r}
portbb <- c(left = min(portland.combined.cat$lon), 
                              bottom = min(portland.combined.cat$lat), 
                              right = max(portland.combined.cat$lon), 
                              top = max(portland.combined.cat$lat))
```


```{r}
portmap <- get_map(portbb, zoom = 11, scale = 2, maptype="terrain",source="google")
```
```{r}
portland.map <- ggmap(portmap) + geom_point(data = portland.combined.cat, aes(x = lon, y = lat,fill=DN_AF1), color = "black",pch=21, size = 3) + theme_minimal() +scale_fill_viridis_c(option = "inferno")+labs(title="Portland",x="Longitude",y="Latitude",fill="Afternoon\nTemp (F)") +theme(plot.title = element_text(size = 14, hjust = .5,face = "bold.italic"))
```

```{r portland-map, fig.height=7, fig.width=7}
portland.map
```


```{r}
portland.map.cat <- ggmap(portmap) + geom_point(data = portland.combined.cat, aes(x = lon, y = lat,fill=Category), color = "black",pch=21, size = 3) + theme_minimal() +scale_fill_manual(name="Category",values=c("#7fcdbb","#fe9929","#DDA0DD"))+labs(title="Portland",x="Longitude",y="Latitude",fill="Conservation\nCategory") +theme(plot.title = element_text(size = 14, hjust = .5,face = "bold.italic"))
```


```{r portland-map-concern, fig.height=7, fig.width=7}
portland.map.cat
```

# Seattle

```{r}
summary <- seattle.combined.cat %>% group_by(city.space,Category) %>% summarize(n=n(),meanAF=mean(DN_AF1,na.rm=TRUE),n_AF_na=sum(is.na(DN_AF1)))
summary
```

```{r}
ggplot(summary,aes(n,city.space,fill=Category))+geom_col(position=position_dodge(),alpha=0.7) +theme_bw() + scale_fill_manual(name="Category",values=c("#7fcdbb","#fe9929","#DDA0DD")) +labs(x="Number of Trees",y="Tree Inventory")
```


## Seattle Map

```{r}
seabb <- c(left = min(seattle.combined.cat$lon), 
                              bottom = min(seattle.combined.cat$lat), 
                              right = max(seattle.combined.cat$lon), 
                              top = max(seattle.combined.cat$lat))
```


```{r}
seamap <- get_map(seabb, zoom = 11, scale = 2, maptype="terrain",source="google")
```

```{r}
seattle.map <- ggmap(seamap) + geom_point(data = seattle.combined.cat, aes(x = lon, y = lat,fill=DN_AF1), color = "black",pch=21, size = 3) + theme_minimal() +scale_fill_viridis_c(option = "inferno")+labs(title="Seattle",x="Longitude",y="Latitude",fill="Afternoon\nTemp (F)") +theme(plot.title = element_text(size = 14, hjust = .5,face = "bold.italic"))
```

```{r seattle-map, fig.height=7, fig.width=7}
seattle.map
```


```{r}
seattle.map.cat <- ggmap(seamap) + geom_point(data = seattle.combined.cat, aes(x = lon, y = lat,fill=Category), color = "black",pch=21, size = 3) + theme_minimal() +scale_fill_manual(name="Category",values=c("#7fcdbb","#fe9929","#DDA0DD"))+labs(title="Seattle",x="Longitude",y="Latitude",fill="Conservation\nCategory") +theme(plot.title = element_text(size = 14, hjust = .5,face = "bold.italic"))
```


```{r seattle-map-concern, fig.height=7, fig.width=7}
seattle.map.cat
```


```{r}
seattle.map <- seattle.map + theme(legend.position = "none",axis.title.x = element_blank())
seattle.map.cat  <- seattle.map.cat + theme(legend.position = "none",plot.title = element_blank())
portland.map <- portland.map +theme(axis.title.x = element_blank(),axis.title.y = element_blank())
portland.map.cat <- portland.map.cat +theme(plot.title = element_blank(),axis.title.y = element_blank())


#+ theme(legend.position = "none",,)
```



```{r combined-maps, fig.height=10, fig.width=10}
(seattle.map | portland.map) / (seattle.map.cat | portland.map.cat)
```

