---
title: "Binomial Distribution Response to EHD Metrics"
author: "Joey Hulbert"
date: "Last Update: `r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

|            |            |            |            |            |
|:----------:|:----------:|:----------:|:----------:|:----------:|
|[Welcome](https://jmhulbert.github.io/redhot)|[Data](https://jmhulbert.github.io/redhot/data)|[Analyses](https://jmhulbert.github.io/redhot/analyses)|[Discussion](https://jmhulbert.github.io/redhot/discussion)|[Maps](https://jmhulbert.github.io/redhot/maps)
|             |           |            |            |            |


```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(corrplot)
library(knitr)
library(kableExtra)
library(gghalves)
library(patchwork)
library(scales)
library(glmmTMB)
library(bbmle)
library(DHARMa)
```


# Purpose

The purpose of this page is to summarize an investigation of **tree health categories** as a *categorical* response to EHD Metrics.

Response variable: category: *healthy, unhealthy*

Possible explanitory variables to explore: 

+ EHD Data (see [V2.0 Report](https://doh.wa.gov/sites/default/files/2022-07/311-011-EHD-Map-Tech-Report_0.pdf?uid=62e46bbc98fad) for more details)
  + Indicators in Pollution Burden
    + Environmental Exposures
      + Diesel exhaust
      + Ozone
      + PM2.5
      + Toxic Release
      + Proximity to heavy traffic roadways
    + Environmental Effects
      + Lead risk from housing
      + Proximity to hazardous waste treatment storage
      + Proximity to superfund sites
      + wastewater discharge
  + Indicators of Population Characteristics
    + Sensitive Populations
      + Death from cardiovascular disease
      + Low birth weight
    + Socioeconomic Factors
      + No high school diploma
      + Unaffordable housing
      + Transportation expense
      + Limited English
      + People living in poverty
      + Race (people of color)
      + Unemployment



# Read Data

```{r}
data <- read.csv('https://raw.githubusercontent.com/jmhulbert/redhot/main/data/urban-data-modified.csv')
```

# Wrangle Data

## Remove dead trees

```{r}
# Remove Dead Trees
data.w.dead <- data 
data <- data %>% filter(field.tree.canopy.symptoms!="Tree is dead") %>% filter(field.dieback.percent<100) %>% droplevels()
```

```{r}
data$binary.tree.canopy.symptoms <- as.factor(data$binary.tree.canopy.symptoms)
levels(data$binary.tree.canopy.symptoms)
```

## Limit Data to data with EHD Data (Washington only)

```{r}
data <- data %>% filter(Environm_1!="")
```

Only 787 observations were included in the analyses of ehd ranks in urban areas because the data was limited to washington observations.

Note there are more observations available in the [WRDM](https://www.inaturalist.org/projects/western-redcedar-dieback-map) for these analyses, but this study only includes trees from areas where urban heat data were available.


# Analyses

## Fit Models


### Model 1 

```{r}
binomial.ehd <- glmmTMB(binary.tree.canopy.symptoms ~ Environm_1 + (1|Area),family=binomial,data=data)
```


#### Summary

```{r}
summary(binomial.ehd)
```


#### Interpretation


The probability of the outcome does not change with changes in the environmental health disparity rank.


### Model 2 

```{r}
binomial.poverty <- glmmTMB(binary.tree.canopy.symptoms ~ Poverty + (1|Area),family=binomial,data=data)
```

#### Summary

```{r}
summary(binomial.poverty)
```


The probability of the outcome does not change with changes in poverty.



