---
title: "Binomial Distribution Response to HOLC Ranks"
author: "Joey Hulbert"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

|            |            |            |            |            |
|:----------:|:----------:|:----------:|:----------:|:----------:|
|[Welcome](https://jmhulbert.github.io/redhot)|[Data](https://jmhulbert.github.io/redhot/data)|[Analyses](https://jmhulbert.github.io/redhot/analyses)|[Discussion](https://jmhulbert.github.io/redhot/discussion)|[Maps](https://jmhulbert.github.io/redhot/maps)
|             |           |            |            |            |


```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(corrplot)
library(knitr)
library(kableExtra)
library(gghalves)
library(patchwork)
library(scales)
library(glmmTMB)
library(bbmle)
library(DHARMa)
```


# Purpose

The purpose of this page is to summarize an investigation of **tree health categories** as a *categorical* response to HOLC grades.

Response variable: category: *healthy, unhealthy*


# Read Data

```{r}
data <- read.csv('https://raw.githubusercontent.com/jmhulbert/redhot/main/data/urban-data-modified.csv')
```

# Wrangle Data

## Remove dead trees

```{r}
# Remove Dead Trees
data.w.dead <- data 
data <- data %>% filter(field.tree.canopy.symptoms!="Tree is dead") %>% filter(field.dieback.percent<100) %>% droplevels()
```

```{r}
data$binary.tree.canopy.symptoms <- as.factor(data$binary.tree.canopy.symptoms)
levels(data$binary.tree.canopy.symptoms)
```

## Limit Data to data with HOLC ranks

```{r}
data <- data %>% filter(grade!="")
```


# Analyses

## Fit Models

```{r}
binomial.holc <- glmmTMB(binary.tree.canopy.symptoms ~ grade + (1|Area),family=binomial,data=data)
```


## Summary

```{r}
summary(binomial.holc)
```


## Interpretation
```{r}
levels(as.factor(data$grade))
```

The probability of the outcome does not change when changing HOLC grades from A to B, C or D.

When accounting for cities as a random effect and when limiting data to the 395 observations with HOLC data.

Interesting



